---
title: "Data Generator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Generator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dineR)
```

<small> _This Vignette covers the built-in Data Generating function in depth._ </small>

The focus of *dineR* lies primarily on differential networks, and the subsequent estimation thereof. However, the structure of data suitable for differential network estimation may not be immediately apparent. As such, this function provides users with an easy, scalable manner to generate appropriate data as they familiarize themselves with the estimation options available within *dineR*

## Getting Started

The _data_generator_ function enables users to generate two multivariate normal samples of varying sizes, whose number of features/dimensions are equal. These two samples are representative of the two experimental conditions under which one would typically determine a differential network. 

Due to the technicalities of differential network estimation, it is not possible to validate the accuracy of estimates obtained through the use of traditional statistical methodologies. _data_generator_ aims to address this by generating data whose covariance matrices, and precision matrices are known a-priori and as such so too is the differential network. This allows users to validate the estimation procedure for their specific experimental criteria. 

## Arguments

The most important arguments available within _data_generator_ are as follows:

```{r}
# Number of observations in sample 1
n_X <- 100 

# Number of observations in sample 2
n_Y <- 150

# Number of features in each of the samples
p <- 50

# The form of the precision covariance matrices
case <- "sparse"

# The seed of the simulation process to ensure reproducibility of results
seed <- 123
```

## Data Generation

Having defined each of the key arguments above, the _data_generator_ function can now be called:

```{r}
data <- data_generator(n_X = n_X, n_Y = n_Y, p = p, case = case, seed = seed)
```

## Outputs

Given the data has now been generated, the data can be "wrangled" for use within the estimation function. 

```{r}
# Extract the first sample
X <- data$X
cat("The number of observations in the first sample is:", nrow(X))
cat("The number of features/dimensions in the first sample is:", ncol(X))

# Extract the second sample
Y <- data$Y
cat("The number of observations in the second sample is:", nrow(Y))
cat("The number of features/dimensions in the second sample is:", ncol(Y))
```

It is also possible to extract the sample covariance matrices for each of the samples:

```{r}
# Extract the first sample's covariance matrix
Sigma_X <- data$Sigma_X

# Extract the second sample's covariance matrix
Sigma_Y <- data$Sigma_Y
```

In addition to the above matrices, the precision matrices can also be extracted:

```{r}
# Extract the first sample's precision matrix
Omega_X <- data$Omega_X

# Extract the second sample's precision matrix
Omega_Y <- data$Omega_Y
```

Lastly, it is possible to extract the differential network for comparison to the estimate:

```{r}
# Extract the differential network
Delta <- data$Delta
```

Having covered each of the key arguments and outputs available within _data_generator_, please kindly note that further details are available within the function documentation as well as the listed reference.
